#Exercise Prediction  
#By Shubham Mahajan  

#Synopsis  
In this report we will explore the different ways people did exercises. The data is taken from groupware. Using machine learning techniques we will train a model using the training data set to accurately find the different classe of excercise the people did using the testing data set.  

##Getting and Splitting the Data  
```{r}
set.seed(100)
library(caret)
train <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")

inTrain <- createDataPartition(y=train$classe, p=0.7, list=F)
train1 <- train[ inTrain, ]
train2 <- train[-inTrain, ]
```

##Cleaning Data  
```{r}

#removing the variables with zero variance
novariance <- nearZeroVar(train1)
train1 <- train1[, -novariance]
train2 <- train2[, -novariance]

# remove the variables that are mostly have the NA value

mostNA <- sapply(train1, function(x) mean(is.na(x))) > 0.95
train1 <- train1[, mostNA==FALSE]
train2 <- train2[, mostNA==FALSE]

# remove any useless variables like name, the number of trial, etc. These are the first five variables

train1 <- train1[, -(1:5)]
train2 <- train2[, -(1:5)]

```

##Training the Model  
```{r cache=TRUE}
# I used random forest for modelling and k fold cross validation, in this   case k=3
library(e1071)
threeFoldCV <- trainControl(method="cv", number=3, verboseIter=F)
fit <- train(classe ~ ., data=train1, method="rf", trControl=threeFoldCV)
fit$finalModel
```

##Predicting on the Test Set in the Training Data  
```{r cache=TRUE}
preds <- predict(fit, newdata=train2)
confusionMatrix(train2$classe, preds)
```

##Predicting on the Overall Test Set  
```{r}
# Since the model above worked well, we can use it on the test set
predictions <- predict(fit, newdata=test)

# show the results
results <- data.frame(problem_id=test$problem_id,predicted=predictions)
print(results)

```


